---
title: "PM566 Final Project"
author: "Xiaoyu"
output:
  pdf_document:
    toc: TRUE
    latex_engine: xelatex
  word_document: 
    toc: TRUE
always_allow_html: yes
editor_options:
  chunk_output_type: console
---

```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)
library(skimr)
library(olsrr)
library(mfp)
# INITALIZE CODE CHUNK OPTIONS
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = FALSE,
  cache = TRUE,
  error = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px",
  class.source = "code-r")
```


```{css, echo = FALSE}
.code-r { /* Code block */
  font-size: 15px;
}
.code-r-small { /* Code block */
  font-size: 10px;
}
```

# Introduction

Diabetes is a complex metabolic syndrome and its involvement in various diseases is manifold with varying manifestations and different clinical symptoms and prognosis. Although being probably the most important risk factor, diabetes is often considered an “accompanying comorbidity” for cardiac or peripheral artery disease, hypertension, or stroke. An interesting study demonstrated that the risk of sudden cardiac death is at least 2-times higher in patients who are diabetic compared with those who are nondiabetic, regardless of the extent of cardiac dysfunction or symptoms of heart failure. In contrast, the risk of nonsudden cardiac death was not signiﬁcantly different between the 2 groups.

Analyzing diabetes is beneficial as it could help prevent tragedic illnesses such as sudden cardiac death, the leading cause of death in the United States. The object of this study is to find out if diabetes risk prediction is possible based on family history and genetic factors along with other risk factors. 

This study is focused on 1) whether Diabetes Pedigree Function is truly associated with diabetes and if so, what risk factors have a significant relationship with it. 2) Creating effective diabetes prediction model. The association between Diabetes Pedigree Function and disease status would reveal to what extent does the genetic factors affect diabetes. The dataset did not separate type 1 diabetes and type 2 diabetes. Regardless of the type of diabetes, however, the dataset has sufficient predictors to evaluate the hypothesis because both types of diabetes are caused by impaired glucose metabolism. Diabetes Pedigree Function, which is the key variable, can be assessed as family history is one of the important risk factors for both types of diabetes. In the analysis, important indicators of diabetes will be evaluated together.


## Data description

- Lipotoxicity (1-17): lipotoxicity is a metabolic syndrome that results from the accumulation of lipid intermediates in non-adipose tissue, leading to cellular dysfunction and death.

- Glucose (mmol/L): blood glucose level obtained by measuring plasma glucose concentration at 2 hours in an oral glucose tolerance test.

- BloodPressure (mm Hg): the pressure of the blood in the circulatory system. 

- SkinThickness (mm): skin thickness is primarily determined by collagen content and is increased in insulin-dependent diabetes mellitus.

- Insulin(µU/ml): insulin is an anabolic hormone that promotes glucose uptake.

- BMI: body mass index (BMI) is a person's weight in kilograms divided by the square of height in meters.

- DiabetesPedigreeFunction (0:1 value generated from familial diabetes history/risk): diabetes pedigree function provides “a synthesis of the diabetes mellitus history in relatives and the genetic relationship of those relatives to the subject.” It generally provides scores of the likelihood of diabetes based on family history. The DPF uses information from parents, grandparents, siblings, aunts and uncles, and first cousins. It provides a measure of the expected genetic influence of affected and unaffected relatives on the subject’s eventual diabetes risk. 

- Age: age of the individual.

- Outcome: diabetes test result (0 = Non-diabetic, 1 = Diabetic).


## Key questions

Is there a significant difference in values of diabetes risk factors for those who have diagnosed with diabetes and those who are not? Is Diabetes Pedigree Function significantly associated with the onset of diabetes and other risk factors? Could these risk factors provide a reliable prediction of individual's diabetes?


<br>

# Methods

The dataset used in this study was acquired from Harvard Dataverse. Among several predictors, Diabetes Pedigree Function was a particularly interesting attribute in the dataset. It provided some data on diabetes history in relatives and the genetic relationship of those relatives to the patient. This measure of genetic influence became the starting point of an idea that the hereditary risk could be used for the risk stratification or prediction of onset of diabetes. 

Reference: P. Rodgers, Judith, 2020, "Diabetes Mellitus “Comorbidity” or an Important Risk Factor?", https://doi.org/10.7910/DVN/JAW6AX, Harvard Dataverse, V2, UNF:6:QOwrJ53n2F5fMC+wb4ADVA== [fileUNF]

Analysis through assessing various plots, tables and graphs was performed to identify association between Diabetes Pedigree Function and diabetes test outcome, including the examination of the effect of lipotoxicity, glucose level, blood pressure, skin thickness, insulin level, BMI, and age. The data was cleaned by replacing extreme values to 'NA's, shortening variable names, and creating new factor variable for better analysis. Skim function from skimr package was used to explore data. Dim, head and tail, summary, and table functions were used to check detailed observations. Age was stratified into four age groups(20-29, 30-39, 40-49, and 50+) for better understanding of relationship with diabetes. Outcome was binomial variable(0, 1) and it was transformed into factor variable (Non-diabetic and Diabetic). For prediction model, logistic regression was performed with grouped smooth method and LOESS smoothing method conducted for linearity assumption evaluation. Influential outliers and model fit was checked and the statistics including accuracy, sensitivity, and specificity were displayed with graphs. ROC curve and area under the curve was measured for the discrimination ability of the model.

```{r}
diabetes <- read_csv("diabetes.csv")
```


## Exploratory Data Analysis

- The total dataset includes 768 observations with 9 vaiables. 

- Lipotoxicity is right skewed and ranged from 0 to 17 with mean of 3.85. 

- Glucose level is not skewed to either side and ranged from 0 to 199 with mean of 121. 

- Blood Pressure is also not skewed and ranged from 0 to 122 with mean of 69.1. 

- Skin Thickness right skewed and ranged from 0 to 99. 

- Insulin level is extremly right skewed and ranged from 0 to 846. 

- BMI is fairly normally distributed and ranged from 0 to 67.1. 

- Diabetes Pedigree Function is right skewed and ranged from 0.078 to 2.42.

- Age is right skewed and ranged from 21 to 81. Right skewed age data displays more than 50% of observations are in the age group of 20 to 29. This data is more focused on young population.

- Outcome has binary results with mean of 0.349 which means 34.9% of all observations have diabetes positive outcome.

```{r}
skim(diabetes)
```

## Data Cleaning

- 0 values in all risk factors are speculated as an absence of the specific test outcome because there were significant difference in values between all 0s and the next minimum values of each variable. All of them are edited to "NA"s. 

```{r}
db <-
  diabetes %>%
  mutate(Lipotoxicity = replace(Lipotoxicity, Lipotoxicity==0, NA))%>%
  mutate(Glucose = replace(Glucose, Glucose==0, NA)) %>%
  mutate(BloodPressure = replace(BloodPressure, BloodPressure==0, NA)) %>%
  mutate(SkinThickness = replace(SkinThickness, SkinThickness==0, NA)) %>%
  mutate(Insulin = replace(Insulin, Insulin==0, NA)) %>%
  mutate(BMI = replace(BMI, BMI==0, NA))
  
```

- Variable names are renamed into lowercase letters with shorter length.

```{r}
db <- 
  db %>%
  rename(
    lip = Lipotoxicity,
    glu = Glucose,
    bp = BloodPressure,
    st = SkinThickness,
    ins = Insulin,
    bmi = BMI,
    dpf = DiabetesPedigreeFunction,
    age = Age,
    outcome = Outcome
    )
```


- Created age groups (20-29, 30-39, 40-49, 50+) to compare proportion of the diabetes by age groups. 

```{r}
db$age_group <- ifelse(db$age < 30, "20-29", 
                         ifelse(db$age >= 30 & db$age < 40, "30-39",
                                ifelse(db$age >= 40 & db$age < 50, "40-49", "50+")))
```

<br>

# Results

## Histogram of Diabetes outcome by age groups

- Age is a significant risk factor for diabetes. As age is a well-known confounder of most diseases, it could also play a role as a confounder when generating a prediction model. Through this histogram, I can confirm that age affects the onset of diabetes.

```{r}
db <-
  db %>%
  mutate(outcome_f = factor(outcome))
db$outcome_f <- if_else(db$outcome_f == 0, "Non-diabetic", "Diabetic")
p0 <- 
  db %>%
  ggplot()+
  geom_bar(mapping = aes(x = age_group, fill = outcome_f))+
  scale_fill_manual(name = "Diabetes", values = c("salmon", "cyan3"))+
  labs(title = "Diabetes Outcome by Age Groups") +
  labs(x = "Age Group", y = "Count")
ggplotly(p0)
```


## Mean Values of Each Risk Factors by Outcome Group
- All of the predictors are showing some differences in mean values by diabetes outcome, meaning that these predictors can be utilized for a prediction model.

- Diabetes test outcome: 0 (Non-diabetic), 1 (Diabetic).

```{r}
db_ot <- 
  db %>% 
  group_by(outcome)
  
db_table <-
  db_ot %>%
  summarise(
    DiabetesPedigreeFunction=mean(dpf, na.rm = TRUE),
    Lipotoxicity=mean(lip, na.rm = TRUE),
    Glucose=mean(glu, na.rm = TRUE),
    BloodPressur=mean(bp, na.rm = TRUE),
    SkinThickness=mean(st, na.rm = TRUE),
    Insulin=mean(ins, na.rm = TRUE),
    BMI=mean(bmi, na.rm = TRUE),
    Age=mean(age, na.rm = TRUE),
    )
knitr::kable(db_table, align= "c", digits = 2, caption = "Mean of Each Risk Factors by Outcome Group")
```
<br>

## Box Plot of each risk factors by outcome {.tabset}
- Among the predictors, Glucose Level, Insulin level, and BMI showed significant differences in mean values by outcome group. 
- Insulin level is a direct indicator of discernment between type 1 and 2 diabetes. This predictor can be treated differently by the type of diabetes. In this dataset, it is presumed that there were more type 2 diabetes in the patient group as the result shows that mean insulin level is higher in the diabetic group. 
- Although test for measuring lipotoxicity is not common, lipotoxicity is showing a meaningful gap between the diabetes and non-diabetes group.
- Skin thickning is a symptom detected from patients with insulin-dependent diabetes mellitus (IDDM). It means this data only applies to type 1 diabetes.
- Blood Pressure does not show significant difference by diabetes outcome.

### Lipotoxicity

```{r}
lip_plot <- ggplot(db_ot, aes(x=outcome_f, y=lip, color=outcome_f, fill=outcome_f)) +
  scale_fill_manual(name = "Diabetes", values = c("salmon", "cyan3")) +
  scale_colour_manual(name = NULL, values = c("black", "black")) +
  labs(title = "Lipotoxicity Level by Diabetes Outcome Group") +
  labs(x = "Diabetes", y = "Lipotoxicity")+
  geom_boxplot()
ggplotly(lip_plot)
```

### Glucose Level

```{r}
glu_plot <- ggplot(db_ot, aes(x=outcome_f, y=glu, color=outcome_f, fill=outcome_f)) +
  scale_fill_manual(name = "Diabetes", values = c("salmon", "cyan3")) +
  scale_colour_manual(name = NULL, values = c("black", "black")) +
  labs(title = "Glucose Level by Diabetes Outcome Group") +
  labs(x = "Diabetes", y = "Glucose (mmol/L)")+
  geom_boxplot()
ggplotly(glu_plot)
```

### Blood Pressure

```{r}
bp_plot <- ggplot(db_ot, aes(x=outcome_f, y=bp, color=outcome_f, fill=outcome_f)) +
  scale_fill_manual(name = "Diabetes", values = c("salmon", "cyan3")) +
  scale_colour_manual(name = NULL, values = c("black", "black")) +
  labs(title = "Blood Pressure by Diabetes Outcome Group") +
  labs(x = "Diabetes", y = "Blood Pressure (mm Hg)")+
  geom_boxplot()
ggplotly(bp_plot)
```

### Skin Thickness

```{r}
st_plot <- ggplot(db_ot, aes(x=outcome_f, y=st, color=outcome_f, fill=outcome_f)) +
  scale_fill_manual(name = "Diabetes", values = c("salmon", "cyan3")) +
  scale_colour_manual(name = NULL, values = c("black", "black")) +
  labs(title = "Skin Thickness by Diabetes Outcome Group") +
  labs(x = "Diabetes", y = "Skin Thickness (mm)")+
  geom_boxplot()
ggplotly(st_plot)
```

### Insulin Level

```{r}
ins_plot <- ggplot(db_ot, aes(x=outcome_f, y=ins, color=outcome_f, fill=outcome_f)) +
  scale_fill_manual(name = "Diabetes", values = c("salmon", "cyan3")) +
  scale_colour_manual(name = NULL, values = c("black", "black")) +
  labs(title = "Insulin Level by Diabetes Outcome Group") +
  labs(x = "Diabetes", y = "Insulin (µU/ml)")+
  geom_boxplot()
ggplotly(ins_plot)
```

### BMI

```{r}
bmi_plot <- ggplot(db_ot, aes(x=outcome_f, y=bmi, color=outcome_f, fill=outcome_f)) +
  scale_fill_manual(name = "Diabetes", values = c("salmon", "cyan3")) +
  scale_colour_manual(name = NULL, values = c("black", "black")) +
  labs(title = "BMI by Diabetes Outcome Group") +
  labs(x = "Diabetes", y = "BMI")+
  geom_boxplot()
ggplotly(bmi_plot)
```


## Plots of Diabetes Outcome by Diabetes Pedigree Function {.tabset}

- The histogram shows that Diabetes Pedigree Function follows Poisson distribution. Poisson regression could be used for further analysis.
- The proportion of diabetic outcome increases over Diabetes Pedigree Fucgion. From dpf 0.24 to 0.48 section, proportion of diabetic vs non-diabetic is approximately 1:2. In 0.56 to 0.72 section, the ratio is 2:3 and in 0.8 and over section, the ratio become close to 1:1.
- Boxplot also shows that there is a significant difference in Diabetes Pedigree Function value between diabetic and non-diabetic group. There is more research needed for the outliers without diabetes.

### Histogram

```{r}
p1 <- ggplot(db, aes(x=dpf, color=outcome_f, fill=outcome_f)) +
  scale_fill_manual(name = "Diabetes", values = c("salmon", "cyan3")) +
  scale_colour_manual(name = NULL, values = c("black", "black")) +
  labs(title = "Diabetes Pedigree Fuction Distribution by Diabetes Outcome") +
  labs(x = "Diabetes Pedigree Function", y = "Count")+
  geom_histogram()
ggplotly(p1)
```

### Boxplot

```{r}
dpf_plot <- ggplot(db_ot, aes(x=outcome_f, y=dpf, color=outcome_f, fill=outcome_f)) +
  scale_fill_manual(name = "Diabetes", values = c("salmon", "cyan3")) +
  scale_colour_manual(name = NULL, values = c("black", "black")) +
  labs(title = "Diabetes Pedigree Fuction by Diabetes Outcome Group") +
  labs(x = "Diabetes", y = "Diabetes Pedigree Function")+
  geom_boxplot()
ggplotly(dpf_plot)
```


## Scatter Plot Graph of Significant Risk Factors vs Diabetes Pedigree Function by Diabetes Outcome Group  {.tabset}

- Glucose level and diabetes pedigree fuction interact well to distinguish diabetic section and non-diabetic section.
- Insulin level is not effective on marking off diabetic section in this scatter plot. However, it could be different when it only applies to the dataset of patients who entirely has one of two types of diabetes.
- BMI is also a good partner of diabetes pedigree function. 
- Upper right side of the lipotoxicity vs DPF graph displays more diabetic outcome as both lipotoxicity and diabetes pedigree function are high.

### Glucose vs DPF

```{r}
db_ot %>% 
  plot_ly(x = ~dpf, y = ~glu, 
          type = 'scatter', mode = 'markers', color = ~outcome_f, colors = "Set1",
          hoverinfo = 'text',
          text = ~paste( paste("Diabetes Outcome:", outcome_f, sep=""), paste(" Diabetes Pedigree Function: ", dpf, sep="") , paste(" Glucose Level: ", glu, sep=""), sep = "<br>")) %>%
  layout(title = "Glucose Level vs Diabetes Pedigree Function by oucome",
                  yaxis = list(title = "Glucose Level"), xaxis = list(title = "Diabetes Pedigree Function"),
         hovermode = "compare")
```

<br>

### Insulin vs DPF

```{r}
db_ot %>% 
  plot_ly(x = ~dpf, y = ~ins, 
          type = 'scatter', mode = 'markers', color = ~outcome_f, colors = "Set1",
          hoverinfo = 'text',
          text = ~paste( paste("Diabetes Outcome:", outcome_f, sep=""), paste(" Diabetes Pedigree Function: ", dpf, sep="") , paste(" Insulin Level: ", ins, sep=""), sep = "<br>")) %>%
  layout(title = "Insulin Level vs Diabetes Pedigree Function by oucome",
                  yaxis = list(title = "Insulin Level"), xaxis = list(title = "Diabetes Pedigree Function"),
         hovermode = "compare")
```

### BMI vs DPF

```{r}
db_ot %>% 
  plot_ly(x = ~dpf, y = ~bmi, 
          type = 'scatter', mode = 'markers', color = ~outcome_f, colors = "Set1",
          hoverinfo = 'text',
          text = ~paste( paste("Diabetes Outcome:", outcome_f, sep=""), paste(" Diabetes Pedigree Function: ", dpf, sep="") , paste(" BMI: ", bmi, sep=""), sep = "<br>")) %>%
  layout(title = "BMI vs Diabetes Pedigree Function by oucome",
                  yaxis = list(title = "BMI"), xaxis = list(title = "Diabetes Pedigree Function"),
         hovermode = "compare")
```

### Lipotoxicity vs DPF

```{r}
db_ot %>% 
  plot_ly(x = ~dpf, y = ~lip, 
          type = 'scatter', mode = 'markers', color = ~outcome_f, colors = "Set1",
          hoverinfo = 'text',
          text = ~paste( paste("Diabetes Outcome:", outcome_f, sep=""), paste(" Diabetes Pedigree Function: ", dpf, sep="") , paste(" Lipotoxicity: ", lip, sep=""), sep = "<br>")) %>%
  layout(title = "Lipotoxicity vs Diabetes Pedigree Function by oucome",
                  yaxis = list(title = "Lipotoxicity"), xaxis = list(title = "Diabetes Pedigree Function"),
         hovermode = "compare")
```

# Conclusion

Overall, most of the predictors displayed differences in mean values when it compares between diabetic and non-diabetic groups. The interesting predictor was diabetes pedigree function because, unlike other risk factors, DPF could be measured by relatives history and genetic data. This is the only predictor that can be obained by external sources other than individual's biological test result. In the analysis, diabetes pedigree function showed its association with the onset of diabetes. Although it is hard to predict the risk of diabetes with diabetes pedigree fuction alone, the analysis showed a possibility of utilization of other risk factors combined with diabetes pedigree funtion in prediction model. In this study, glucose level and BMI provided evidence of significant association with onset of diabetes in conjunction of diabetes pedigree function. Insulin level and skin thickness could also be the part of prediction model when it focuses on one of two types of diabetes.

## Prediction Model

```{r}
library(mfp)
library(ROCit)
library(interactions)
source("plot_resid_lev_logistic.R")
source("logit_plot.R")
source("group_smooth.R")
```

### Dataset Information
- The dataset includes 768 observations (500 non-diabetic, 268 diabetic). Training dataset accounts 70% of randomly selected observations and test dataset accounts rest of 30% of observations.

```{r}
with(db,
     gmodels::CrossTable(outcome))
db_pred <-
db %>%
  mutate(training = sample(0:1, nrow(.),prob=c(.3, .7), replace=T))
db_train <- db_pred %>% filter(training == 1)
db_test  <- db_pred %>% filter(training == 0)
```

### Univariate Analysis

- Before I build a preliminary main effects model, I performed univariable analaysis. All the variables in the dataset was analysed as there was no unappropriate predictor based on the table and plots assessment. Multivariable fractional polynomial was used to obtain a best fit model of each variable. 

```{r}
mfp::mfp(outcome ~ fp(dpf), family = binomial, data = db_train)
group_smooth("dpf", "outcome", db_train)
m1 <- glm(outcome ~ dpf, family = binomial, data = db_train)
summary(m1)
confint(m1)
m1 %>%
  ggplot(aes(x=dpf,y=outcome)) +
  geom_smooth()+
  geom_point()
mfp::mfp(outcome ~ fp(lip), family = binomial, data = db_train)
logit_plot("lip", "outcome", db_train)
m2 <- glm(outcome ~ lip, family = binomial, data = db_train) 
summary(m2)
confint(m2)
mfp::mfp(outcome ~ fp(glu), family = binomial, data = db_train)
logit_plot("glu", "outcome", db_train)
m3 <- glm(outcome ~ glu, family = binomial, data = db_train)
summary(m3)
confint(m3)
mfp::mfp(outcome ~ fp(bp), family = binomial, data = db_train)
logit_plot("bp", "outcome", db_train)
m4 <- glm(outcome ~ bp, family = binomial, data = db_train)
summary(m4)
confint(m4)
mfp::mfp(outcome ~ fp(st), family = binomial, data = db_train)
logit_plot("st", "outcome", db_train)
m5 <- glm(outcome ~ st, family = binomial, data = db_train)
summary(m5)
confint(m5
        )
mfp::mfp(outcome ~ fp(ins), family = binomial, data = db_train)
logit_plot("ins", "outcome", db_train)
m6 <- glm(outcome ~ ins, family = binomial, data = db_train)
summary(m6)
confint(m6)
mfp::mfp(outcome ~ fp(bmi), family = binomial, data = db_train)
logit_plot("bmi", "outcome", db_train)
m7 <- glm(outcome ~ bmi, family = binomial, data = db_train)
summary(m7)
confint(m7)
mfp::mfp(outcome ~ age, family = binomial, data = db_train)
group_smooth("age", "outcome", db_train)
logit_plot("age", "outcome", db_train)
m8 <- glm(outcome ~ age, family = binomial, data = db_train)
summary(m8)
confint(m8)
```


```{r, eval=FALSE}
subset_outcome_prelim <-
  glmulti::glmulti(outcome ~ dpf + lip + glu + bp + st + ins + bmi + age, 
          data = db_train,
          level=1, family = binomial, crit="aicc", confsetsize=128)
best_subset_outcome_prelim <- 
  summary(subset_outcome_prelim)$bestmodel %>% glm(., data = db_train, family = binomial)
summary(best_subset_outcome_prelim)
```

```{r, eval=FALSE}
subset_outcome_prelim2 <-
  glmulti::glmulti(outcome ~ dpf + lip + glu + ins + bmi, 
          data = db_train,
          level=2, family = binomial, crit="aicc", confsetsize=128)
best_subset_outcome_prelim2 <- 
  summary(subset_outcome_prelim2)$bestmodel %>% glm(., data = db_train, family = binomial)
summary(best_subset_outcome_prelim2)
best_outcome_model <- glm(outcome ~ dpf + lip + glu + ins + bmi + ins*dpf + ins*lip + dpf*bmi, family=binomial, data=db_train) 
summary(best_outcome_model)
sim_slopes(best_outcome_model, pred = ins, modx = dpf, johnson_neyman = FALSE)
sim_slopes(best_outcome_model, pred = ins, modx = lip, johnson_neyman = FALSE)
sim_slopes(best_outcome_model, pred = dpf, modx = bmi, johnson_neyman = FALSE)
db_train %>%
  ggplot(aes(x=ins, y=dpf)) +
  geom_point() +
  geom_smooth()
```


```{r}
glm(outcome ~ dpf + lip + glu + ins + bmi + ins*dpf, family=binomial, data=db_train) %>% summary()
glm(outcome ~ dpf + lip + glu + ins + bmi + age + ins*dpf, family=binomial, data=db_train) %>% summary()
```

- The parameter estimate for lipotoxicity changed by 68% after including age into the model. It appears that age confounds the relationship between lipotoxicity and diabetes outcome.
```{r}
glm(outcome ~ lip, family = binomial, data = db_train) %>% summary()
glm(outcome ~ lip + age, family = binomial, data = db_train) %>% summary()
```

```{r}
final_model <- glm(outcome ~ dpf + lip + glu + ins + bmi + age + ins*dpf, family=binomial, data=db_train) 
```

```{r}
ResourceSelection::hoslem.test(final_model$y, fitted(final_model), g=20)
library(LogisticDx)
plot_resid_lev_logistic(final_model)
```

```{r}
DescTools::Conf(final_model, pos = 1)
final_model.p <-
  tibble(
    pred_p = final_model$fitted.values,
    y = final_model$y
  )
# Accuracy
final_model.roc <- 
  ROCit::measureit(score = final_model$fitted.values, 
                   class = final_model$y,
                   measure = c("ACC", "SENS", "SPEC"))
tibble(
  Cutoff = final_model.roc$Cutoff,
  ACC = final_model.roc$ACC
) %>%
ggplot(aes(x = Cutoff, y = ACC)) +
  geom_point() +
  geom_line()
# ROC Curve
tibble(
  Cutoff = final_model.roc$Cutoff,
  SENS = final_model.roc$SENS,
  SPEC = final_model.roc$SPEC
) %>%
  pivot_longer(., cols = c("SENS", "SPEC"), values_to = "value", names_to = "metric") %>%
  ggplot(aes(x = Cutoff, y = value, color = metric)) +
  geom_point() + 
  geom_line()
  
roc_empirical <- 
  rocit(score = final_model$fitted.values, class = final_model$y)
plot(roc_empirical, YIndex = F)
summary(roc_empirical)
ciAUC(roc_empirical)
OptimalCutpoints::optimal.cutpoints(X = "pred_p", status = "y", 
                  data = data.frame(final_model.p), 
                  methods = c("Youden", "MaxSpSe", "MaxProdSpSe"), tag.healthy = 0)
```


```{r}
# Extend to the testing data set
final_model_test <- glm(outcome ~ dpf + lip + glu + ins + bmi + age + ins*dpf, family=binomial, data=db_test) 
DescTools::Conf(final_model_test, pos = 1)
test_model.p <-
  tibble(
    pred_p = predict(final_model, newdata = db_test, type = "response"),
    y = db_test$outcome
  )
skim(test_model.p)
test_model.roc <- 
  ROCit::measureit(score = predict(final_model, newdata = db_test, type = "response"), 
                   class = db_test$outcome,
                   measure = c("ACC", "SENS", "SPEC"))
tibble(
  Cutoff = test_model.roc$Cutoff,
  SENS = test_model.roc$SENS,
  SPEC = test_model.roc$SPEC
) %>%
  pivot_longer(., cols = c("SENS", "SPEC"), values_to = "value", names_to = "metric") %>%
  ggplot(aes(x = Cutoff, y = value, color = metric)) +
  geom_point() + 
  geom_line()
test_roc_empirical <-
  rocit(score = predict(final_model, newdata = db_test, type = "response"), 
        class = db_test$outcome)
plot(test_roc_empirical, YIndex = F)
summary(test_roc_empirical)
ciAUC(test_roc_empirical)
```